Towards human SuperEEG

### 2018-10-07 paper/main.tex
finishing up results, backing up progress on most informative locations analysis [WIP]

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/methods}
  \caption{\textbf{Methods overview.}  \textbf{A.  Electrode
      locations.}  Each dot reflects the location of a single
    electrode in dataset 1, colored according to 7 factor labels (see
    Panel D for details). One patient's electrode locations are
    highlighted in black and the to-be-reconstructed recording
    location is highlighted in red. \textbf{B. Radial basis function
      (RBF).}  Each electrode contributed by the patient (black)
    weights on the full set of locations under consideration (all dots
    in Panel A, defined as $\bar{R}$ in the text).  The weights fall
    off with positional distance (in MNI space) according to an RBF.
    \textbf{C. Per-patient correlation matrices.}  After computing the
    pairwise correlations between the recordings from each patient's
    electrodes, we use RBF-weighted averages to estimate correlations
    between all locations in $\bar{R}$.  We obtain an estimated
    full-brain correlation matrix using each patient's
    data. \textbf{D.  Combined correlation matrix.}  We estimate a
    single full-brain correlation matrix by averaging the
    patient-specific correlation matrices.  We sort the resulting
    correlation matrix based on 7 factor labels obtained from k-means
    clustering~\cite{YeoEtal11}). \textbf{E.  Reconstructing activity
      throughout the brain.}  Given the observed activity from the
    patient's electrodes and the estimated correlation matrix (Panel
    D), we can compute a maximum likelihood estimate of the voltage
    trace at any location in $\bar{R}$.  An example reconstruction (at
    the red dot in Panel A) is shown in red, and the actual recording
    at that location is highlighted above in blue.}
  \label{fig:methods}
\end{figure}

@@

Although both datasets we examined provide good full-brain coverage
(when considering data from every patient; e.g.\
Fig.~\ref{fig:corrmap}B, D), electrodes are not placed uniformly
throughout the brain.  For example, electrodes are more likely to be
implanted in regions like the medial temporal lobe (MTL), and are
rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the XXX~mm$^3$ voxel
MNI152 brain, we computed the proportion of electrodes contained
within a 20 MNI unit radius sphere centered on that voxel.  We defined
the \textit{density} at that location as this proportion.  Across
Datasets 1 and 2, the electrode placement densities were similar
(correlation by voxel: $r = 0.57, p < 10^{-10}$).  We wondered whether
regions with good covererage might be associated with better
reconstruction accuracy (e.g. Fig.~\ref{fig:corrmap}B, D indicate that
many electrodes in the MTL have relatively high reconstruction
accuracy, and occiptal electrodes tend to have relatively low
reconstruction accuracy).  To test whether this held more generally
across the entire brain, for each dataset we computed the electrode
placement density for each electrode from each patient (using the
proportion of \textit{other} patients' electrodes within 20 MNI units
of the given electrode).  We then correlated these density values with
the across-patient reconstruction accuracies for each electrode.
Contrary to our expectation, rather than positive correlations, we
found weak (but reliable) negative correlations between reconstruction
accuracy and density for both datasets (Dataset 1: $r = XXX, p = XXX$;
Dataset 2: $r = -0.16, p < 10^{-10}$).  This indicates that the
reconstruction accuracies we observed are not driven solely by
sampling density, but rather may also reflect higher order properties
of neural dynamics such as functional correlations between distant
voxels~\citep{BetzEtal17b}.

\textbf{JRM STOPPED HERE}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/density}
  \caption{\textbf{Sampling density and reconstruction quality.}
    \textbf{A. \& B. } The glass brain maps show sampling density by voxel location for dataset 1 and dataset 2. \textbf{C.}
      Correlation of sampling density by voxel location for dataset 1 vs. dataset 2.}
  \label{fig:density}
\end{figure}

In addition to exploring how reconstruction quality varies with
location, we also wondered whether there might be effects of electrode
placements on reconstruction quality.  For example, are there
particular implantation locations that yield especially high
reconstruction accuracies at other locations throughout the brain? To gain insights into this questions, we computed the average reconstruction correlation for each patient, then computed the average patient reconstruction correlation for any patients who had electrodes within a 20 MNI unit diameter sphere centered on each voxel location. The resulting maps highlight the locations of implanted electrodes from patients whose
reconstructions were especially accurate  (Fig.~\ref{fig:informap}A and B). We found that the most informative locations were consistent across datasets which lends support to the notion that different electrode location are more informative about activity across patients (Fig.~\ref{fig:informap}C); $r = 0.22, p < 10^{-10}$).  The locations in dark red might therefore be good candidate
implantation targets for neurosurgeons and neurologists who wish to
use SuperEEG to reconstruct full-brain electrophysiological signals.
The above findings, that one can infer brain activity throughout a person's brain 
using recordings from a limited number of locations from that person's brain in conjunction with recordings from other people's brains, have deep
implications for the structure of brain data.  The first implication
is that the correlational structure of different people's brain data
is largely preserved across individuals. Despite recent evidence that different people have stable but reliably different resting state connectome~\cite{FinnEtal15}, our results suggest that the correlational structure of different people's brain data is preserved enough across individuals to provide meaningful information.

>>>>

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/methods}
  \caption{\textbf{Methods overview.}  \textbf{A.~Electrode
      locations.}  Each dot reflects the location of a single
    electrode implanted in the brain of a Dataset 1 patient.  A
    held-out recording location from one patient is indicated in red,
    and the patient's remaining electrodes are indicated in black.
    The electrodes from the remaining patients are colored by
    $k$-means cluster (computed using the full-brain correlation model
    shown in Panel D).  \textbf{B.~Radial basis function kernel.}
    Each electrode contributed by the patient (black) weights on the
    full set of locations under consideration (all dots in Panel A,
    defined as $\bar{R}$ in the text).  The weights fall off with
    positional distance (in MNI space) according to an RBF.
    \textbf{C.~Per-patient correlation matrices.}  After computing the
    pairwise correlations between the recordings from each patient's
    electrodes, we use RBF-weighted averages to estimate correlations
    between all locations in $\bar{R}$.  We obtain an estimated
    full-brain correlation matrix using each patient's
    data. \textbf{D.~Merged correlation model.}  We combine the
    per-patient correlation matrices (Panel C) to obtain a single
    full-brain correlation model that captures information contributed
    by every patient.  Here we have sorted the rows and columns to
    reflect $k$-means clustering labels~\citep[using
    $k$=7;][]{YeoEtal11}, whereby we grouped locations based on their
    correlations with the rest of the brain (i.e., rows of the matrix
    displayed in the panel).  The boundaries denote the cluster
    groups.  The rows and columns of Panel C have been sorted using
    teh Panel D-derived cluster labels.  \textbf{E.~Reconstructing activity
      throughout the brain.}  Given the observed recordings from the
    given patient (shown in black; held-out recording is shown in
    blue), along with a full-brain correlation model (Panel D), we
    use Equation~\ref{eqn:reconstruction} to reconstruct the most
    probable activity at the held-out location (red).}
  \label{fig:methods}
\end{figure}

@@

Although both datasets we examined provide good full-brain coverage
(when considering data from every patient; e.g.\
Fig.~\ref{fig:corrmap}B, D), electrodes are not placed uniformly
throughout the brain.  For example, electrodes are more likely to be
implanted in regions like the medial temporal lobe (MTL), and are
rarely implanted in occipital cortex (Fig.~\ref{fig:density}A, B).
Separately for each dataset, for each voxel in the 4~mm$^3$ voxel
MNI152 brain, we computed the proportion of electrodes contained
within a 20 MNI unit radius sphere centered on that voxel.  We defined
the \textit{density} at that location as this proportion.  Across
Datasets 1 and 2, the electrode placement densities were similar
(correlation by voxel: $r = 0.57, p < 10^{-10}$).  We wondered whether
regions with good covererage might be associated with better
reconstruction accuracy (e.g. Fig.~\ref{fig:corrmap}B, D indicate that
many electrodes in the MTL have relatively high reconstruction
accuracy, and occiptal electrodes tend to have relatively low
reconstruction accuracy).  To test whether this held more generally
across the entire brain, for each dataset we computed the electrode
placement density for each electrode from each patient (using the
proportion of \textit{other} patients' electrodes within 20 MNI units
of the given electrode).  We then correlated these density values with
the across-patient reconstruction accuracies for each electrode.
Contrary to our expectation, rather than positive correlations, we
found weak (but reliable) negative correlations between reconstruction
accuracy and density for both datasets (Dataset 1: $r = XXX, p = XXX$;
Dataset 2: $r = -0.16, p < 10^{-10}$).  This indicates that the
reconstruction accuracies we observed are not driven solely by
sampling density, but rather may also reflect higher order properties
of neural dynamics such as functional correlations between distant
voxels~\citep{BetzEtal17b}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/density}
  \caption{\textbf{Electrode sampling density by location.}
    \textbf{A.~Electrode sampling density by voxel in Dataset 1.} Each
    voxel is colored by the proportion of total electrodes in the
    dataset that are
    located within a 20~MNI unit radius sphere centered on the given
    voxel.  \textbf{B.~Electrode sampling density by voxel in Dataset
      2.}  This panel displays the sampling density map for Dataset 2,
    in the same format as Panel A.  \textbf{C.~Correspondence in
      sampling density by voxel across Datasets 1 and 2.}  The
    two-dimensional histogram displays the by-voxel densities in the
    two Datasets, and the one-dimensional histograms display the
    proportions of voxels in each dataset with the given density
    value.  The correlation reported in the panel is across voxels in
    the 4~mm$^3$ MNI brain.}
  \label{fig:density}
\end{figure}

In neurosurgical applications where one wishes to infer full-brain
activity patterns, can our framework yield insights into where the
electrodes should be placed?  A basic assumption of our approach (and
of most prior ECoG work) is that electrode recordings are most
informative about the neural activity near the recording surface of
the electrode.  But if we consider that activity patterns throughout
the brain are meaningfully correlated, are there particular
implantation locations that, if present in a patient's brain, yield
especially high reconstruction accuracies throughout the rest of the
brain?  For example, one might hypothesize that brain structures that
are heavily interconnected with many other structures could be more
informative about full-brain activity patterns than comparatively
isolated structures.

To gain insights into whether particular electrode locations might be
especially informative, we first computed the average reconstruction
accuracy across all of each patient's electrodes (using the
across-patients cross validation test; black histograms in
Fig.~\ref{fig:corrmap}A and C).  We labeled each patient's electrodes
in each dataset with the average reconstruction accuracy for that
patient.  In other words, we assigned every electrode from each
patient the same value, reflecting how well the activity patterns at
those electrodes were reconstructed on average.  Next, for each voxel
in the 4~mm$^3$ MNI brain, we computed the average value across any
electrode (from any patient) that came within 20 MNI units of that
voxel's center.  Effectively, we computed the average reconstruction
accuracy across any patients with electrodes near each voxel, where
the avearges were weighted to reflect patients who had more electrodes
implanted near that location.

\textbf{JRM STOPPED HERE}
% - report most-informative locations (i.e. which brain structures?)
% - can we say anything bout whether these most-informative regions are heavily
% interconnected, e.g. perhaps from the resting state literature?

To gain insights into this question, we
computed the average reconstruction accuracy for each patient, then
computed the average patient reconstruction correlation for any
patients who had electrodes within a 20 MNI unit diameter sphere
centered on each voxel location. The resulting maps highlight the
locations of implanted electrodes from patients whose reconstructions
were especially accurate (Fig.~\ref{fig:informap}A and B). We found
that the most informative locations were consistent across datasets
which lends support to the notion that different electrode location
are more informative about activity across patients
(Fig.~\ref{fig:informap}C); $r = 0.22, p < 10^{-10}$).  The locations
in dark red might therefore be good candidate implantation targets for
neurosurgeons and neurologists who wish to use SuperEEG to reconstruct
full-brain electrophysiological signals.  The above findings, that one
can infer brain activity throughout a person's brain using recordings
from a limited number of locations from that person's brain in
conjunction with recordings from other people's brains, have deep
implications for the structure of brain data.  The first implication
is that the correlational structure of different people's brain data
is largely preserved across individuals. Despite recent evidence that
different people have stable but reliably different resting state
connectome~\cite{FinnEtal15}, our results suggest that the
correlational structure of different people's brain data is preserved
enough across individuals to provide meaningful information.
